---
title: "Exploring Hotel Review Data from Trip Advisor"
output: html_document
---

I wanted to explore hotel review data for a few hotels from [Trip Advisor](http://www.tripadvisor.com). I have described scraping data for 3 hotels [here](http://notesofdabbler.github.io/201408_hotelReview/scrapeTripAdvisor.html). The hotels considered are (disclaimer: choice of hotels is random):

* [J. W. Marriott, Indianapolis](http://www.tripadvisor.com/Hotel_Review-g37209-d1762915-Reviews-JW_Marriott_Indianapolis-Indianapolis_Indiana.html)
* [Conrad, Indianapolis](http://www.tripadvisor.com/Hotel_Review-g37209-d576787-Reviews-Conrad_Indianapolis-Indianapolis_Indiana.html)
* [Hampton Inn Indanapolis Northwest - Park 100, Indianapolis](http://www.tripadvisor.com/Hotel_Review-g37209-d1236866-Reviews-Hampton_Inn_Indianapolis_Northwest_Park_100-Indianapolis_Indiana.html)

## Loading and Reviewing Data

```{r loaddata,warning=FALSE,tidy=FALSE}
# set working directory
setwd("~/notesofdabbler/githubfolder/blog_notesofdabbler/hotelReview/")

library(dplyr)
library(lubridate)
library(ggplot2)
library(tm)
library(scales)
library(topicmodels)

# load data for a hotel
# currently 3 datasets available from scraping trip advisor data
#    1. J W Marriott, Indianapolis (label: jwmarriott)
#    2. Hampton Inn Indianapolis Northwest -100 (label: hamptoninn)
#    3. Conrad Indianapolis (label: conrad)
#

hotellist=c("jwmarriott","conrad","hamptoninn")


# load hotel review data for each hotel and return the following for each hotel
# * Data frame with hotel reviews
# * Top 3 records
# * Number of reviews in dataset
# * frequency of different ratings
dfrating.l=lapply(hotellist,
                  function(x) {
                      filenm=paste("dfrating_",x,".Rda",sep="")
                      load(filenm)
                      return(list(dfrating=dfrating,
                                  top3records=head(dfrating,3),
                                  numrevs=nrow(dfrating),
                                  freqRating=table(dfrating$ratingnum)))
                                         })
names(dfrating.l)=hotellist

dfrating.l[["jwmarriott"]]$top3records
dfrating.l[["jwmarriott"]]$numrevs
dfrating.l[["jwmarriott"]]$freqRating

dfrating.l[["conrad"]]$top3records
dfrating.l[["conrad"]]$numrevs
dfrating.l[["conrad"]]$freqRating

dfrating.l[["hamptoninn"]]$top3records
dfrating.l[["hamptoninn"]]$numrevs
dfrating.l[["hamptoninn"]]$freqRating
```

## Ratings Over Time

These 3 hotels are listed among the top 5 hotels in Indianapolis. So most of the ratings are 4 or 5 for these. Next, I have plotted % of reviews with a given rating over time

```{r revbymon,warning=FALSE,tidy=FALSE}
# function to get number of reviews and % of reviews with a rating
getrev.bymon=function(hotel){
  
  # get review data
  dfrating=dfrating.l[[hotel]]$dfrating

  # create a month label
  dfrating$yrmon=floor_date(dfrating$ratingdt,"month")
  
  # create sequence of months
  yrmon=unique(dfrating$yrmon)
  yrmonseq=seq(min(yrmon),max(yrmon),by="months")

  # yr-month and rating combinations
  yrmon.rating=expand.grid(yrmon=yrmonseq,ratingnum=c(1,2,3,4,5))
  
  # get % of reviews for each rating by month
  dfrating.bymon=dfrating%>%group_by(yrmon,ratingnum)%>%summarize(count=n())
  dfrating.bymon.agg=dfrating.bymon%>%group_by(yrmon)%>%summarize(countfull=sum(count))
  dfrating.bymon=merge(dfrating.bymon,dfrating.bymon.agg,c("yrmon"))
  dfrating.bymon$pctrating=dfrating.bymon$count/dfrating.bymon$countfull
  
  
  dfrating.bymon=merge(yrmon.rating,dfrating.bymon,by=c("yrmon","ratingnum"),all.x=TRUE)
  dfrating.bymon$pctrating[is.na(dfrating.bymon$pctrating)]=0
  dfrating.bymon$count[is.na(dfrating.bymon$count)]=0
  dfrating.bymon$countfull[is.na(dfrating.bymon$countfull)]=0
  
  dfrating.bymon$hotel=hotel
  
  # get number of reviews by month
  dfrating.bymon.revs=dfrating.bymon%>%group_by(yrmon)%>%summarize(numrevs=sum(count))
  dfrating.bymon.revs$hotel=hotel
  
  return(list(dfrating.bymon=dfrating.bymon,dfrating.bymon.revs=dfrating.bymon.revs))

}

dfrating.bymon.l=lapply(hotellist,function(x) getrev.bymon(x))


# plot number of reviews by year-month
dfrating.bymon.revs=do.call(rbind,lapply(dfrating.bymon.l,function(x) x$dfrating.bymon.revs))
p=ggplot(dfrating.bymon.revs,aes(x=yrmon,y=numrevs))+geom_line()+facet_grid(hotel~.)
p=p+xlab("")+ylab("# of reviews")
p=p+theme_bw()
p

# plot % of reviews for each rating by year-month
dfrating.bymon=do.call(rbind,lapply(dfrating.bymon.l,function(x) x$dfrating.bymon))
p=ggplot(dfrating.bymon,aes(x=yrmon,y=pctrating,color=factor(ratingnum)))+geom_line(size=1.1)
p=p+facet_grid(hotel~.)
p=p+xlab("")+ylab("% of Ratings")
p=p+scale_y_continuous(breaks=seq(0,1,0.1),labels=percent)+scale_color_discrete(name="# stars")
p=p+theme_bw()
p

```

The more choppy areas of % of ratings plot correspond to periods with very less reviews.

## Words in Top Quote for a Review

I wanted to check out the top words occuring in top quotes associated with review for each star rating. Though I show the top words for each start rating, we should remember that the frequency of lower start ratings in data is low.

```{r,warning=FALSE,tidy=FALSE}
# Explore top level quotes for each rating

# function to get document-term matrix from hotel review data for a given hotel
getDTM=function(dftxt){

  # code adapted from http://www.rdatamining.com/examples/text-mining
  
  txtcorpus=Corpus(VectorSource(dftxt))
  #inspect(txtcorpus[1:5])
  
  txtcorpus.cl=tm_map(txtcorpus,tolower)
  txtcorpus.cl=tm_map(txtcorpus.cl,removePunctuation)
  txtcorpus.cl=tm_map(txtcorpus.cl,removeNumbers)
  
  mystopwords=c(stopwords("english"),"hotel","staff","room","rooms","indianapolis","marriott","conference",
                "convention","indy","downtown","hampton","stay","stayed","inn","conrad")
  txtcorpus.cl=tm_map(txtcorpus.cl,removeWords,mystopwords)
  #dictCorpus=txtcorpus.cl
  #txtcorpus.cl=tm_map(txtcorpus.cl,stemDocument)
  #txtcorpus.cl=tm_map(txtcorpus.cl,stemCompletion,dictionary=dictCorpus)
  
  dtm=DocumentTermMatrix(txtcorpus.cl)
  
  dtm.m=as.matrix(dtm)
  
  return(dtm.m)
}

getTopTerms=function(hotel){

  # get review data
  dfrating=dfrating.l[[hotel]]$dfrating
  
  minrating=min(dfrating$ratingnum)
  maxrating=max(dfrating$ratingnum)
  tfreq.l=as.list(rep(NA,maxrating-minrating+1))
  
  # of frequent words to retain
  numterms=20
  
  rating=maxrating+1
  
  for(i in 1:(maxrating-minrating+1)){
    
    rating=rating-1
    #sprintf("Processing data for %s stars",rating)
    
    dftxt=dfrating$topquote[dfrating$ratingnum==rating]
    
    dtm.m=getDTM(dftxt)
    
    tfreq=colSums(dtm.m)
    tfreq.l[[i]]=names(sort(tfreq,decreasing=TRUE)[1:numterms])
  }
  
  topTerms=do.call(cbind,tfreq.l)
  colnames(topTerms)=paste(seq(maxrating,minrating)," star")
  
  return(list(dtm.m=dtm.m,topTerms=topTerms))
  
}

topTerms.l=lapply(hotellist,function(x) getTopTerms(x))
names(topTerms.l)=hotellist

topTerms.l[["jwmarriott"]]$topTerms
topTerms.l[["conrad"]]$topTerms
topTerms.l[["hamptoninn"]]$topTerms

```

## Exploring Full Review Text

Next, I wanted to explore the full text review to see if there are themes. Right now, I am not too familiar with topic model packages (on my near term todo list). For now, I took a more simpler approach of just using k-means clustering of words using term document term matrix find cluster of words. Here I just picked 5 clusters and haven't checked what the right number of clusters are.

```{r,warning=FALSE,tidy=FALSE}
# Further investigation of high star rating using full reviews for a hotel

# functiont to cluster words from term document matrix of full reviews for a hotel

getClust=function(hotel){
  
  # load hotel full review data
  dfrating=dfrating.l[[hotel]]$dfrating
  
  dftxt=dfrating$fullrev[dfrating$ratingnum>=4]
  dtm.m=getDTM(dftxt)
  
  # clustering of words to detect themes/topics
  
  set.seed(1234)
  txtclust=kmeans(t(dtm.m),5)
  
  # size of clusters
  txtclust$size
  
  # within and total sum of squares
  txtclust$totss
  txtclust$withinss
  
  # get list of frequent terms in each cluster
  clustTerms=as.list(rep(NA,5))
  
  termlist=colnames(dtm.m)
  for(i in 1:5){
    termlist.filt=termlist[txtclust$cluster == i]
    tfreq=colSums(dtm.m)
    tfreq.filt=sort(tfreq[termlist.filt],decreasing=TRUE)
    clustTerms[[i]]=names(tfreq.filt[1:20])  
  }
  clust.topic=do.call(cbind,clustTerms)
  clust.topic[is.na(clust.topic)]=""
  colnames(clust.topic)=c("cluster 1","cluster 2","cluster 3","cluster 4","cluster 5")
  
  # print list of frequent terms in each cluster
  return(list(txtclust=txtclust,clust.topic=clust.topic))
  
}

clust.topic=lapply(hotellist,function(x) getClust(x))
names(clust.topic)=hotellist

clust.topic[["jwmarriott"]]$clust.topic

clust.topic[["conrad"]]$clust.topic

clust.topic[["hamptoninn"]]$clust.topic

```

I was finding it hard to assign themes that differentiate clusters. Perhaps I need to do a better job filtering words or play around more with parameters for clustering or check out algorithms for topic models. Nevertheless, this was a fun exercise.

## Session Info
All analysis was done in RStudio 0.98.994

```{r}
sessionInfo()
```