
```{r setup, include=FALSE}
knitr::opts_chunk$set(warnings=FALSE,tidy=FALSE)
```

# Explore Kaggle User Profile

I extracted the user information from [kaggle](http://www.kaggle.com) for about top 500 ranked kaggle users and created an exploratory plots of best rank for a user by country and number of users by country.

## Scraping User Information

The [following]("http://www.kaggle.com/users") site has users listed in the order of rank over multiple pages. About 13 pages cover users with ranks upto 520. I extracted the information on user name, user rank, number of competitions and location country. Not all of the users had listed the country and so there were some unknown locations in the data.

```{r}
# set working directory
setwd("~/notesofdabbler/githubfolder/blog_notesofdabbler/exploreKaggle/")

# load libraries
library(rvest)
library(stringr)
library(dplyr)
library(googleVis)
op <- options(gvis.plot.tag='chart')

library(ggplot2)

# Get url of pages that contain users
# 13 pages correspond to about top 500 ranks

url = "http://www.kaggle.com/users"
urllist = paste(url,seq(2,13),sep="?page=")
urllist = c(url,urllist)
head(urllist)

# Function to parse user information from each page
getURL = function(url){
    # get list of users
    users = html(url) %>% html_nodes(".users-list")
    # get user names
    usrname = users %>% html_nodes(".profilelink") %>% html_text()
    # get user rank
    usrrnk = users %>% html_nodes(".rank") %>% html_text()
    # get number of competitions a user participated in 
    usrnumcomp = users %>% html_nodes(".comps") %>% html_text()
    # get user country 
    usrloc = users %>% html_nodes("li") %>% html_text()
    usrloc2 = sapply(as.list(usrloc),function(x) {
        xsplit = strsplit(x,"\r\n")[[1]]
        xtrim = str_trim(xsplit)
        xnoblank = xtrim[xtrim != ""]
        xloc = xnoblank[length(xnoblank)]
        return(xloc)
      })
    # combine into a dataframe
    usrdf = data.frame(usrname,usrrnk,usrnumcomp,usrloc2,stringsAsFactors = FALSE)
    return(usrdf)
}

# compile user information 
usrdf = list()
length(usrdf) = length(urllist)

for(i in 1:length(urllist)){
  usrdf[[i]] = getURL(urllist[i])
}

usrdf2 = rbind_all(usrdf)
head(usrdf2)

```

The user country field needed some manual cleaning to match the country name needed for googleVis geo chart. I exported the country list in the extracted Kaggle user data to a csv file, manually appended the country names that are needed for googleVis chart and imported it back.

```{r}
# clean up location field 
# currently logic of extracting location was looking for last entry in a vector
# if location is unknown, the last entry corresponds to number of competitions
# if location field has number of competitions, then it is set to unknown
tmpqc = usrdf2[grepl("competition",usrdf2[["usrloc2"]]),]
usrdf2[["usrloc2"]][grepl("competition",usrdf2[["usrloc2"]])] = "unknown"

# Convert user rank to a numeric value
usrdf2["usrrnk2"] = as.numeric(gsub("[^(0-9)]","",usrdf2[["usrrnk"]]))
# Convert number of competitions to a numeric value
usrdf2["usrnumcomp2"] = as.numeric(gsub("[^(0-9)]","",usrdf2[["usrnumcomp"]]))

# Get list of countries
cntrycnt = usrdf2 %>% group_by(usrloc2) %>% summarize(cntrycnt = n()) %>% arrange(desc(cntrycnt))
cntrylist = cntrycnt$usrloc2

# write country list to a csv file for manual cleaning
# Needed to manually assign country name that googleVis chart recognizes
#write.csv(cntrylist,"cntrylist.csv")

# read in the cleaned country list
cntrylist_cleaned = read.csv("cntrylist_cleaned.csv",sep=",")

# merge cleaned country list
usrdf3 = merge(usrdf2,cntrylist_cleaned,by.x=c("usrloc2"),by.y=c("cntryName"))

head(usrdf3)
```

## Analyze and Visualize Data

I summarize the data by country to get the following:

* Number of users
* Averge rank
* Best rank
* Competitions per user

```{r}
# Summary by country
#  - number of users
#  - average rank of users
#  - best rank of users
#  - competitions per user
cntrycnt3 = usrdf3 %>% group_by(cntryName2)%>%
                  summarize(cntrycnt = n(),avgrnk = mean(usrrnk2),bestrnk = min(usrrnk2),
                            totcomp = sum(usrnumcomp2)) %>% 
                  arrange(bestrnk)
cntrycnt3["compperusr"] = cntrycnt3[["totcomp"]]/cntrycnt3[["cntrycnt"]]
head(cntrycnt3)
```

The plot below shows the best rank in each country

```{r,results='asis'}
# plot user data (best rank by country) in a geochart
pltdf = cntrycnt3 %>% select(-totcomp)
names(pltdf) = c("country","NumberOfUsers","AverageRank","BestRank","CompetitionsPerUser")
G = gvisGeoChart(pltdf, locationvar = "country",
                        colorvar = "BestRank",
                        options=list(height=600,width=600))
T = gvisTable(pltdf,options = list(height=600,width=600),
              format=list(AverageRank = "#",
                          CompetitionsPerUser = "#.#"))

GT = gvisMerge(G,T,horizontal = TRUE)
plot(GT)
```

The plot below shows the number of users by country (caveat: this is just based on top 500 ranked users)

```{r,results='asis'}
# Plot number of users by country
G2 = gvisGeoChart(pltdf, locationvar = "country",
                  colorvar = "NumberOfUsers",
                  options=list(height=600,width=600))
plot(G2)
```

## Session Info

All analysis was done with RStudio 0.98.1062. 

```{r}
sessionInfo()
```